### 2021-04-26 sections/queueing_model/general-markov-model/main.tex
Add general Markov-chain picture

\begin{figure}[H]
    \centering
    \scalebox{.8}
    {
        \begin{tikzpicture}[-, node distance = 0.9cm, auto, every node/.style={scale=0.7}]

            % Markov chain variables
            \tikzmath{
                let \initdist = 0.5cm;
                let \altdist = 1.2cm;
                let \minsz = 1.6cm;
            }

            % S_1 and S_2 rectangles
            \tikzmath{
                let \leftOne = -0.8;
                let \rightOne = 2.7;
                let \upOne = 0.8;
                let \downOne = -2.7;
                let \leftTwo = 2.8;
                let \rightTwo = 13;
                let \upTwo = -2.95;
                let \downTwo = -16.4;
            }

            % General case variables
            \tikzmath{
                let \GCsmallx = 8.3;
                let \GCsmally = -9.5;
                let \GCbigx = 4.1;
                let \GCbigy = -11.8;
            }

            % Rectangle for S1
            \draw[ultra thin, dashed] (\leftOne, \downOne) -- (\leftOne, \upOne);
            \draw[ultra thin, dashed] (\leftOne, \upOne) -- (\rightOne, \upOne);
            \draw[ultra thin, dashed] (\rightOne, \upOne) -- node 
            {\Huge{\( \quad S_1 \)}}(\rightOne, \downOne);
            \draw[ultra thin, dashed] (\rightOne, \downOne) -- (\leftOne, \downOne);

            % Rectangle for S2
            \draw[ultra thin, dashed] (\leftTwo, \downTwo) -- node 
            {\Huge{\( S_2 \quad \)}}(\leftTwo, \upTwo);
            \draw[ultra thin, dashed] (\leftTwo, \upTwo) -- (\rightTwo, \upTwo);
            \draw[ultra thin, dashed] (\rightTwo, \upTwo) -- (\rightTwo, \downTwo);
            \draw[ultra thin, dashed] (\rightTwo, \downTwo) -- (\leftTwo, \downTwo);

            % Small square of general case
            \draw [thick] (\GCsmallx, \GCsmally) -- node {} 
            (\GCsmallx + 0.4, \GCsmally);
            \draw [thick] (\GCsmallx + 0.4, \GCsmally) -- node {} 
            (\GCsmallx + 0.4, \GCsmally - 0.4);
            \draw [thick] (\GCsmallx + 0.4, \GCsmally - 0.4) -- node {} 
            (\GCsmallx, \GCsmally - 0.4);
            \draw [thick] (\GCsmallx, \GCsmally - 0.4) -- node {} 
            (\GCsmallx, \GCsmally);

            % Dashed lines to from small square to big one 
            \draw [ultra thin] (\GCsmallx, \GCsmally) -- node {} 
            (\GCbigx, \GCbigy);
            \draw [ultra thin] (\GCsmallx + 0.4, \GCsmally) -- node {} 
            (\GCbigx + 4, \GCbigy);
            \draw [ultra thin] (\GCsmallx, \GCsmally - 0.4) -- node {} (7, \GCbigy);
            \draw [ultra thin] (\GCsmallx + 0.4, \GCsmally - 0.4) -- node {} 
            (\GCbigx + 4, \GCbigy - 4);

            % Big Square of general case
            \draw [ultra thick] (\GCbigx, \GCbigy) -- node {} (\GCbigx + 4, \GCbigy);
            \draw [ultra thick] (\GCbigx + 4, \GCbigy) -- node {} 
            (\GCbigx + 4, \GCbigy - 4);
            \draw [ultra thick] (\GCbigx + 4, \GCbigy - 4) -- node {General Case} 
            (\GCbigx, \GCbigy - 4);
            \draw [ultra thick] (\GCbigx, \GCbigy - 4) -- node {} (\GCbigx, \GCbigy);

            % First Line
            \node[state, minimum size=1.5cm] (zero) {(0,0)};
            \node[state, node distance = \initdist, minimum size=\minsz, below right=of zero] 
            (one) {(0,1)};
            \node[draw=none, node distance = \initdist, minimum size=\minsz, below right=of one] 
            (two) {\textbf{\( \ddots \)}};
            \node[state, node distance = \initdist, minimum size=\minsz, below right=of two] 
            (three) {(0,T)};
            \node[state, node distance = \altdist, minimum size=\minsz, right=of three] 
            (four) {(0,T+1)};
            \node[draw=none, node distance = \altdist, minimum size=\minsz, right=of four] 
            (five) {\textbf{\dots}};
            \node[state, minimum size=\minsz, right=of five] (six) {(0,C)};
            \node[draw=none, minimum size=\minsz, right=of six] (seven) {\textbf{\dots}};

            % Second Line
            \node[state, minimum size=\minsz, below=of three] (three_one) {(1,T)};
            \node[state, minimum size=\minsz, below=of four] (four_one) {(1,T+1)};
            \node[draw=none, minimum size=\minsz, below=of five] (five_one) {\textbf{\dots}};
            \node[state, minimum size=\minsz, right=of five_one] (six_one) {(1,C)};
            \node[draw=none, minimum size=\minsz, right=of six_one] (seven_one) {\textbf{\dots}};

            % Third Line
            \node[state, minimum size=\minsz, below=of three_one] (three_two) {(2,T)};
            \node[state, minimum size=\minsz, below=of four_one] (four_two) {(2,T+1)};
            \node[draw=none, minimum size=\minsz, below=of five_one] (five_two) 
            {\textbf{\dots}};
            \node[state, minimum size=\minsz, right=of five_two] (six_two) {(2,C)};
            \node[draw=none, minimum size=\minsz, right=of six_two] (seven_two) 
            {\textbf{\dots}};

            % Fourth line
            \node[draw=none, node distance = \altdist, minimum size=\minsz, below=of three_two] 
            (three_three) {\textbf{\vdots}};
            \node[draw=none, node distance = \altdist, minimum size=\minsz, below=of four_two] 
            (four_three) {\textbf{\vdots}};
            \node[draw=none, node distance = 2cm, minimum size=\minsz, below=of five_two] 
            (five_three) {};
            \node[draw=none, node distance = \altdist, minimum size=\minsz, below=of six_two] 
            (six_three) {\textbf{\vdots}};

            % Fifth line
            \node[draw=none, node distance = 0.3cm, minimum size=\minsz, below=of four_three] 
            (general_case_up) {};
            \node[state, node distance = \altdist, minimum size=\minsz, below=of general_case_up] 
            (general_case_mid) {\( (u_i, v_i) \)};

            \node[draw=none, node distance = \altdist, minimum size=\minsz, below=of general_case_mid] 
            (general_case_down) {};
            \node[draw=none, node distance = \altdist, minimum size=\minsz, left=of general_case_mid] 
            (general_case_left) {};
            \node[draw=none, node distance = \altdist, minimum size=\minsz, right=of general_case_mid] 
            (general_case_right) {};

            \draw[every loop]
                % First Horizontal Edges
                (zero) edge[bend left] node {\( \Lambda \)} (one)
                (one) edge[bend left] node {\( \mu \)} (zero)
                (one) edge[bend left] node {\( \Lambda \)} (two)
                (two) edge[bend left] node {\( 2 \mu \)} (one)
                (two) edge[bend left] node {\( \Lambda \)} (three)
                (three) edge[bend left] node {\( T \mu \)} (two)
                (three) edge[bend left] node {\( \lambda_1 \)} (four)
                (four) edge[bend left] node {\( (T+1) \mu \)} (three)
                (four) edge[bend left] node {\( \lambda_1 \)} (five)
                (five) edge[bend left] node {\( (T+2) \mu \)} (four)
                (five) edge[bend left] node {\( \lambda_1 \)} (six)
                (six) edge[bend left] node {\( C\mu \)} (five)
                (six) edge[bend left] node {\( \lambda_1 \)} (seven)
                (seven) edge[bend left] node {\( C\mu \)} (six)

                % Second Horizontal Edges
                (three_one) edge[bend left] node {\( \lambda_1 \)} (four_one)
                (four_one) edge[bend left] node {\( (T+1) \mu \)} (three_one)
                (four_one) edge[bend left] node {\( \lambda_1 \)} (five_one)
                (five_one) edge[bend left] node {\( (T+2) \mu \)} (four_one)
                (five_one) edge[bend left] node {\( \lambda_1 \)} (six_one)
                (six_one) edge[bend left] node {\( C\mu \)} (five_one)
                (six_one) edge[bend left] node {\( \lambda_1 \)} (seven_one)
                (seven_one) edge[bend left] node {\( C\mu \)} (six_one)

                % Third Horizontal Edges
                (three_two) edge[bend left] node {\( \lambda_1 \)} (four_two)
                (four_two) edge[bend left] node [below] {\( (T+1) \mu \)} (three_two)
                (four_two) edge[bend left] node {\( \lambda_1 \)} (five_two)
                (five_two) edge[bend left] node {\( (T+2) \mu \)} (four_two)
                (five_two) edge[bend left] node {\( \lambda_1 \)} (six_two)
                (six_two) edge[bend left] node {\( C\mu \)} (five_two)
                (six_two) edge[bend left] node {\( \lambda_1 \)} (seven_two)
                (seven_two) edge[bend left] node {\( C\mu \)} (six_two)

                % First Vertical Edges
                (three) edge[bend left] node {\( \lambda_2 \)} (three_one)
                (three_one) edge[bend left] node {\( T \mu \)} (three)
                (three_one) edge[bend left] node {\( \lambda_2 \)} (three_two)
                (three_two) edge[bend left] node {\( T\mu \)} (three_one)
                (three_two) edge[bend left] node {\( \lambda_2 \)} (three_three)
                (three_three) edge[bend left] node {\( T\mu \)} (three_two)

                % Second Vertical Edges
                (four) edge node {\( \lambda_2 \)} (four_one)
                (four_one) edge node {\( \lambda_2 \)} (four_two)
                (four_two) edge node {\( \lambda_2 \)} (four_three)

                % Fourth Vertical Edges
                (six) edge node {\( \lambda_2 \)} (six_one)
                (six_one) edge node {\( \lambda_2 \)} (six_two)
                (six_two) edge node {\( \lambda_2 \)} (six_three)

                % General Case
                (general_case_left) edge[bend left] node {\( \lambda_1 \)} (general_case_mid)
                (general_case_mid) edge[bend left] node {\( v_i \mu \)} (general_case_left)
                (general_case_right) edge[bend left] node {\( (v_i +1) \mu \)} (general_case_mid)
                (general_case_mid) edge[bend left] node {\( \lambda_1 \)} (general_case_right)
                % (five_three) edge node {\( \lambda_2 \)} (general_case_mid)
                (general_case_up) edge node {\( \lambda_2 \)} (general_case_mid)
                (general_case_mid) edge node {\( \lambda_2 \)} (general_case_down)
                ;
        \end{tikzpicture}
    }
    \caption{Markov chain} 
    \label{fig:general-markov_model}
\end{figure}