### 2021-05-01 sections/queueing_model/main.tex
Apply more siggestions from review
- Markov chains intro
- State space definition
- Replace q_{i,j} with Q_{ij}
- Minor changes

\section{A queueing model with 2 consecutive buffer centres}

% TODO: If this is to be one paper maybe the first two paragraphs do not need to
% include so much details or even better omit some of the details in the game
% overview section.
In this section, a more in-depth explanation of the queueing model shown in 
figure \ref{fig:diagram_of_queueing_system} will be given.
This is a queuing model that consists of two waiting spaces, one for each type
of patient.

The model consists of two types of individuals; class 1 and class 2.
Class 1 individuals arrive instantly at waiting zone 1 and proceed to wait to
receive their service. 
Class 2 individuals arrive at waiting zone 2 and wait there until they are 
allowed to move to waiting zone 1. 
They are allowed to proceed only when the number of 
individuals in waiting zone 1 \textbf{and} in service is less than a 
pre-determined threshold \(T\).
When the number of individuals is equal to or exceeds this threshold, all second type individuals that arrive will remain 
\textit{``blocked''} in waiting zone 1 until the number of people in the 
system is reduced below \(T\). 
This is shown diagrammatically in figure \ref{fig:diagram_of_queueing_system}.
The parameters of the described queueing model are:

\begin{itemize}
    \item \(\lambda_i\): The arrival rate of individuals of type \(i\in\{1, 2\}\)
    \item \(\mu\): The service rate for individuals receiving service
    \item \(C\): The number of servers
    \item \(T\): The threshold at which individuals of the second type are blocked
\end{itemize}

Under the assumption that all rates (arrival and service) are Markovian the
queuing system corresponds to a Markov chain~\cite{kemeny1976markov}.
The states of the Markov chain are denoted by \((u,v)\) where:

\begin{itemize}
    \item \(u\) is the number of individuals blocked
    \item \(v\) is the number of individuals either in waiting zone 1 or in the
    service centre
\end{itemize}

We denote the state space of the Markov chain as  \(S=S(T)\) which can be 
written as the disjoint union~\ref{eq:definition_of_S_as_disjoint_union}.

\begin{align}
    S(T) =& S_1(T) \cup S_2(T) \text{ where:} \nonumber \\
    S_1(T) =& \left\{(0, v)\in\mathbb{N}_0^2 \; | \; v < T \right\} 
    \label{eq:definition_of_S_as_disjoint_union} \\
    S_2(T) =& \{(u, v)\in\mathbb{N}_0^2 \; | \; v \geq T \} \nonumber
\end{align}

The transition matrix \(Q\) of the Markov chain consists of the transition rates
between the numerous states of the model. Every entry \( Q_{ij} = 
Q_{(u_i, v_i),(u_j, v_j)} \) represents the transition rate from state 
\( i = (u_i, v_i) \) to state \( j = (u_j , v_j) \) for all 
\( (u_i, v_i), (u_j, v_j) \in S \).
The entries of \(Q\) can be calculated using the state-mapping function 
described in \ref{eq:markov_transition_rate}:

\begin{equation} \label{eq:markov_transition_rate}
    Q_{ij} = 
    \begin{cases}
        \Lambda, & \textbf{if } (u_i, v_i) - (u_j, v_j) = (0,-1) \textbf{ and } 
        v_i < \text{t} \\
        \lambda_1, & \textbf{if } (u_i, v_i) - (u_j, v_j) = (0,-1) 
        \textbf{ and } v_i \geq \text{t} \\
        \lambda_2, & \textbf{if } (u_i, v_i) - (u_j, v_j) = (-1,0) \\
        v_i \mu, & \textbf{if } (u_i, v_i) - (u_j, v_j) = (0,1) \textbf{ and } 
        v_i \leq C \textbf{ or} \\ & \hspace{0.37cm}(u_i, v_i) - (u_j, v_j) = 
        (1,0) \textbf{ and } v_i = T \leq C \\
        C \mu, & \textbf{if } (u_i, v_i) - (u_j, v_j) = (0,1) \textbf{ and } 
        v_i > C 
        \textbf{ or} \\ & \hspace{0.37cm}(u_i, v_i) - (u_j, v_j) = (1,0) 
        \textbf{ and } v_i = T > C\\
        -\sum_{j=1}^{|Q|}{Q_{ij}} & \textbf{if } i = j \\
        0, & \textbf{otherwise}
    \end{cases}
\end{equation}

Note that \(\Lambda\) here denotes the overall arrival rate in the model by both 
classes of individuals (i.e. \(\Lambda = \lambda_1 + \lambda_2\)). 
A visualisation of how the transition rates relate to the states of the model 
can be seen in the general Markov chain model shown in figure 
\ref{fig:general-markov-model}.

\input{sections/queueing_model/general_markov_model/main.tex}

In order to acquire an exact solution of the problem a slight adjustment needs 
to be considered. 
The problem defined above assumes no upper boundary to the number of individuals 
that can wait for service or for the ones that are blocked in the buffer centre. 
Therefore, a different state space \( \tilde S \) needs to be constructed where 
\( \tilde S \subseteq S \) and there is a maximum allowed number of people \(N\) 
that can be in the system and a maximum allowed number of people \(M\) that can
be blocked in the buffer centre:

\begin{equation}
    \tilde S = \left\{ (u, v) \in S\;| u \leq M, v\leq N \right\}
\end{equation}

\input{sections/queueing_model/subsections/performance_measures/main.tex}

\input{sections/queueing_model/subsections/example/main.tex}