### 2021-05-20 sections/queueing_model/subsections/performance_measures/waiting_time/main.tex
Restructure waiting time formula section
The section is now structured in the
following way:

- Formula 1
- Formula 2
- Formula 3
- Defining necessary parameters

\subsubsection{Waiting time} \label{sec:waiting_time}

Waiting time is the amount of time that individuals from either type wait in 
waiting zone 1 so that they can receive their service. 
For a given set of parameters there are three different performance measures 
around the mean waiting time that can be calculated; the mean waiting time of
type 1 individuals:

\begin{equation} \label{eq:closed_form_waiting_type_1}
    W^{(1)} = \frac{\sum_{\substack{(u,v) \, \in S_A^{(1)} \\ v \geq C}} 
    \frac{1}{C \mu} \times (v-C+1) \times \pi(u,v)}{\sum_{(u,v) \, 
    \in S_A^{(1)}} \pi(u,v)}
\end{equation}

The mean waiting time of type 2 individuals:

\begin{equation}\label{eq:closed_form_waiting_type_2}
    W^{(2)} = \frac{\sum_{\substack{(u,v) \, \in S_A^{(2)} \\ min(v,T) \geq C}} 
    \frac{1}{C \mu} \times (\min(v+1,T)-C) \times \pi(u,v)}{\sum_{(u,v) \, 
    \in S_A^{(2)}} \pi(u,v)}
\end{equation}

The overall mean waiting time:

\begin{equation}\label{eq:overall_waiting_time}
    W = \frac{\lambda_1 P(L'_1)}{\lambda_2 P(L'_2) + \lambda_1 P(L'_1)} W^{(1)} 
    + \frac{\lambda_2 P(L'_2)}{\lambda_2 P(L'_2) + \lambda_1 P(L'_1)} W^{(2)}
\end{equation}

Here \(S_A^{(1)}\) and \(S_A^{(2)}\) are the set of accepting states for type
1 and type 2 individuals. These are the set of states that the model is able
to accept a specific type of individuals.

\begin{equation}\label{eq:accepting_states_type_1}
    S_A^{(1)} = \{(u, v) \in S \; | \; v < N \}
\end{equation}

\begin{equation}\label{eq:accepting_states_type_2}
    S_A^{(2)}=
    \begin{cases}
        \{(u, v) \in S \; | \; u < M \}, & \textbf{if } T \leq N\\
        \{(u, v) \in S \; | \; v < N \}, & \textbf{otherwise}
    \end{cases}
\end{equation}

Equation \ref{eq:overall_waiting_time} makes use of the proportion of type 1 
and type 2 individuals that are not lost to the system. These probabilities are 
given by \(P(L'_1)\) and \(P(L'_2)\) where:

\begin{equation}
    P(L'_1) = \sum_{(u,v) \, \in S_A^{(1)}} \pi(u,v) \hspace{2cm}
    P(L'_2) = \sum_{(u,v) \, \in S_A^{(2)}} \pi(u,v)
\end{equation}