### 2021-05-21 sections/queueing_model/subsections/performance_measures/blocking_time/main.tex
Restructure blocking time formula
Remove most of text that focuses on the "how" part of the formula.
What's left now is just the formula and everything that's necessary to
use it.

\subsubsection{Blocking time}

% TODO: Possibly replace the contents of this section with the ones for the 
% closed form formula of the blocking time
% Currently the direct approach one is shown

Unlike the waiting time, the blocking time is only calculated for individuals of
the second type.  
That is because individuals of the first type cannot be blocked. 
Thus, one only needs to consider the pathway of type 2 individuals to get the 
mean blocking time of the system. 
The mean blocking time can by calculated using:

% TODO: Fix this equation by using the arriving state 
\begin{equation}\label{eq:algebraic_blocking_time}
    B = \frac{\sum_{(u,v) \in S_A^{(2)}} \pi_{(u,v)} \; 
    b(u,v)}{\sum_{(u,v) \in S_A^{(2)}} \pi_{(u,v)}}
\end{equation}

Here \(S_A^{(2)}\) is the set of accepting states of type 2 individuals (defined
in (\ref{eq:accepting_states_type_2})) and \(b(u,v)\) is the mean time that an 
individual will be blocked when the system is at state \((u,v)\). 
For all the states of the system \(b(u,v)\) is given by:

\begin{equation}\label{eq:general_blocking_time_at_each_state}
    b(u,v) = 
    \begin{cases} 
        0, & \textbf{if } (u,v) \notin S_b \\
        c(u,v) + b(u - 1, v), & \textbf{if } v = N = T\\
        c(u,v) + b(u, v-1), & \textbf{if } v = N \neq T \\
        c(u,v) + p_s(u,v) b(u-1, v) + p_a(u,v) b(u, v+1), & \textbf{if } u > 0 
        \textbf{ and } \vspace{-0.2cm} \\ 
        & \quad v = T \\
        c(u,v) + p_s(u,v) b(u, v-1) + p_a(u,v) b(u, v+1), & \textbf{otherwise} \\
    \end{cases}
\end{equation}

Note that \(S_b\) is defined as the set of states where individuals can be
blocked and is given by:

\begin{equation} \label{eq:set_of_blocking_states}
    S_b = \{(u,v) \in S \; | \; u > 0\}
\end{equation}

Additionally, \(c(u,v)\) is the mean sojourn time for each state and \(p_s\) 
and \(p_a\) are the probabilities that the next event to occur will be a service completion or an arrival of a type 1 individual:

\begin{equation}\label{eq:sojourn_blocking_time}
    c(u,v) = 
    \begin{cases}
        \frac{1}{\min(v,C) \mu}, & \text{if } v = N\\
        \frac{1}{\lambda_1 + \min(v,C) \mu}, & \text{otherwise}
    \end{cases}
\end{equation}

\begin{equation}\label{eq:probs_of_service_and_arrival}
    p_s(u,v) = \frac{\min(v,C)\mu}{\lambda_1 + \min(v,C)\mu}, \qquad
    p_a(u,v) = \frac{\lambda_1}{\lambda_1 + \min(v,C)\mu}
\end{equation}

% TODO: These might possibly need to be moved to the appendix 
The system of equations produced by (\ref{eq:general_blocking_time_at_each_state}) can be solved by considering the linear system \(Zx=y\), where:

\begin{equation}\label{eq:general_algebaric_approach_blocking_time}
    \scalebox{0.73}{\(
        Z = 
        \begin{pmatrix}
            -1 & p_a & 0 & \dots & 0 & 0 & 0 & 0 & 0 & \dots & 0 & 0 \\ %(1,T)
            p_s & -1 & p_a & \dots & 0 & 0 & 0 & 0 & 0 & \dots & 0 & 0 \\ %(1,T+1)
            0 & p_s & -1 & \dots & 0 & 0 & 0 & 0 & 0 & \dots & 0 & 0 \\ %(1,T+2)
            \vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & 
            \vdots & \vdots & \ddots & \vdots & \vdots \\ 
            0 & 0 & 0 & \dots & 1 & -1 & 0 & 0 & 0 & \dots & 0 & 0 \\ %(1,N)
            p_s & 0 & 0 & \dots & 0 & 0 & -1 & p_a & 0 & \dots & 0 & 0 \\ %(2,T)
            0 & 0 & 0 & \dots & 0 & 0 & p_s & -1 & p_a & \dots & 0 & 0 \\ %(2,T+1)
            \vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & 
            \vdots & \vdots & \ddots & \vdots & \vdots \\ 
            0 & 0 & 0 & \dots & 0 & 0 & 0 & 0 & 0 & \dots & 1 & -1 \\ %(M,N)
        \end{pmatrix},
        x = 
        \begin{pmatrix}
            b(1,T) \\
            b(1,T+1) \\
            b(1,T+2) \\
            \vdots \\
            b(1,N) \\
            b(2,T) \\
            b(2,T+1) \\
            \vdots \\
            b(M,N) \\
        \end{pmatrix}, 
        y= 
        \begin{pmatrix}
            -c(1,T) \\
            -c(1,T+1) \\
            -c(1,T+2) \\
            \vdots \\
            -c(1,N) \\
            -c(2,T) \\
            -c(2,T+1) \\
            \vdots \\
            -c(M,N) \\
        \end{pmatrix}
    \)}
\end{equation}

Matrix \(Z\) can be alternatively defined using \(Z_{ij}\) defined in equation 
(\ref{eq:general_mapping_function_of_blocking_matrix}) where \(i\) and \(j\) 
represent states \((u_i, v_i), (u_j, v_j) \in S_b\).

\begin{equation}\label{eq:general_mapping_function_of_blocking_matrix}
    Z_{ij} = 
    \begin{cases}
        p_a, & \textbf{if } j = i + 1 \textbf{ and } v_i \neq N \\
        \\
        p_s, & \textbf{if } j = i - 1 \textbf{ and } v_i \neq N, v_i \neq T \\
             & \textbf{or } j = i - N + T \textbf{ and } u_i \geq 2,\,v_i = T \\
        \\
        1, & \textbf{if } j = i - 1 \textbf{ and } v_i = N \\
        \\
        -1, & \textbf{if } i = j \\
        \\
        0, & \textbf{otherwise} \\
    \end{cases}
\end{equation}