### 2021-05-21 sections/appendix/proportion_within_target_formula/main.tex

\section{Mean blocking time} \label{sec:appendix_mean_proportion}