### 2021-05-21 sections/appendix/waiting_time_formula/main.tex

\section{Mean waiting time}\label{sec:appendix_mean_waiting}