### 2021-05-25 sections/appendix/waiting_time_formula/main.tex
Add comparisons between Markov and DES
Include violinplots of waiting time, blocking time and proportion
within target time. The violinplots show the differences between
Markov values and simulated values.

\section{Mean waiting time}\label{sec:appendix_mean_waiting}

% TODO: Include recursive formula of waiting time