### 2021-05-26 sections/queueing_model/subsections/performance_measures/waiting_time/main.tex
Add new violinplots
Include violin plots of performance measures over
different values of lambda_2 for type 1, type 2 and
overall individuals

\subsubsection{Waiting time} \label{sec:waiting_time}

Waiting time is the amount of time that individuals from either type wait in 
waiting zone 1 so that they can receive their service. 
For a given set of parameters there are three different performance measures 
around the mean waiting time that can be calculated; the mean waiting time of
type 1 individuals:

\begin{equation} \label{eq:closed_form_waiting_type_1}
    W^{(1)} = \frac{\sum_{\substack{(u,v) \, \in S_A^{(1)} \\ v \geq C}} 
    \frac{1}{C \mu} \times (v-C+1) \times \pi(u,v)}{\sum_{(u,v) \, 
    \in S_A^{(1)}} \pi(u,v)}
\end{equation}

The mean waiting time of type 2 individuals:

\begin{equation}\label{eq:closed_form_waiting_type_2}
    W^{(2)} = \frac{\sum_{\substack{(u,v) \, \in S_A^{(2)} \\ min(v,T) \geq C}} 
    \frac{1}{C \mu} \times (\min(v+1,T)-C) \times \pi(u,v)}{\sum_{(u,v) \, 
    \in S_A^{(2)}} \pi(u,v)}
\end{equation}

The overall mean waiting time:

\begin{equation}\label{eq:overall_waiting_time}
    W = \frac{\lambda_1 P_{L'_1}}{\lambda_2 P_{L'_2} + \lambda_1 P_{L'_1}} W^{(1)} 
    + \frac{\lambda_2 P_{L'_2}}{\lambda_2 P_{L'_2} + \lambda_1 P_{L'_1}} W^{(2)}
\end{equation}

Here \(S_A^{(1)}\) and \(S_A^{(2)}\) are the set of accepting states for type
1 and type 2 individuals. These are the set of states that the model is able
to accept a specific type of individuals.

\begin{equation}\label{eq:accepting_states_type_1}
    S_A^{(1)} = \{(u, v) \in S \; | \; v < N \}
\end{equation}

\begin{equation}\label{eq:accepting_states_type_2}
    S_A^{(2)}=
    \begin{cases}
        \{(u, v) \in S \; | \; u < M \}, & \textbf{if } T \leq N\\
        \{(u, v) \in S \; | \; v < N \}, & \textbf{otherwise}
    \end{cases}
\end{equation}

Equation \ref{eq:overall_waiting_time} makes use of the proportion of type 1 
and type 2 individuals that are not lost to the system. These probabilities are 
given by \(P_{L'_1}\) and \(P_{L'_2}\) where:

\begin{equation}\label{eq:proportion_of_accepting_individuals}
    P_{L'_1} = \sum_{(u,v) \, \in S_A^{(1)}} \pi(u,v) \hspace{2cm}
    P_{L'_2} = \sum_{(u,v) \, \in S_A^{(2)}} \pi(u,v)
\end{equation}

Appendix \ref{sec:appendix_mean_waiting} gives more details on the recursive
formula that equations (\ref{eq:closed_form_waiting_type_1}),
(\ref{eq:closed_form_waiting_type_2}) and (\ref{eq:overall_waiting_time})
originate from.

Figures \ref{fig:markov_vs_des_waiting_time_comparison_1},
\ref{fig:markov_vs_des_waiting_time_comparison_2} and
\ref{fig:markov_vs_des_waiting_time_comparison_overall} show a comparison 
between the calculated mean waiting time using Markov chains and the simulated 
waiting time using discrete event simulation over a range of values of 
\(\lambda_2\).
The simulation was ran 100 times and the recorded mean waiting time at each 
iteration is used to populate the violin plots.

\begin{figure}[H]
    \centering
    \includegraphics[width=.8\textwidth]{imgs/waiting_time_comparison/waiting_1.pdf}
    \caption{
        Comparison of mean waiting time for type 1 individuals between values 
        obtained from the Markov chain formulas and values obtained from 
        simulation.
    }
    \label{fig:markov_vs_des_waiting_time_comparison_1}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.8\textwidth]{imgs/waiting_time_comparison/waiting_2.pdf}
    \caption{
        Comparison of mean waiting time for type 2 individuals between values 
        obtained from the Markov chain formulas and values obtained from 
        simulation.
    }
    \label{fig:markov_vs_des_waiting_time_comparison_2}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.8\textwidth]{imgs/waiting_time_comparison/waiting_overall.pdf}
    \caption{
        Comparison of mean waiting time for both types of individuals between 
        values obtained from the Markov chain formulas and values obtained from 
        simulation.
    }
    \label{fig:markov_vs_des_waiting_time_comparison_overall}
\end{figure}

% TODO: Possibly add another example that will be used in section 5