### 2021-05-26 sections/queueing_model/proportion_within_target/main.tex

\subsection{Proportion of individuals within target}

Another performance measure that is taken into consideration is the proportion 
of individuals whose waiting and service times lie within a specified 
time target \(t\).
Similar to section \ref{sec:waiting_time} three formulas are needed for this 
performance measure.

The proportion of type 1 individuals within a time target:

\begin{equation}\label{eq:proportion_within_target_type_1}
    P(X^{(1)} < t) = \frac{\sum_{(u,v) \in S_A^{(1)}} P(X_{(u,v)}^{(1)} < t) 
    \pi_{u,v} }{\sum_{(u,v) \in S_A^{(1)}} \pi_{u,v}}
\end{equation}

The proportion of type 2 individuals within a time target:

\begin{equation}\label{eq:proportion_within_target_type_2}
    P(X^{(2)} < t) = \frac{\sum_{(u,v) \in S_A^{(2)}} P(X_{(u,v)}^{(2)} < t) 
    \pi_{u,v} }{\sum_{(u,v) \in S_A^{(2)}} \pi_{u,v}}
\end{equation}

The overall proportion individuals within a time target (where \(P_{L'_1}\) and 
\(P_{L'_1}\) are defined in (\ref{eq:proportion_of_accepting_individuals})):

\begin{align}
    P(X < t) &= \frac{\lambda_1 P_{L'_1}}{\lambda_2 P_{L'_2}+\lambda_1 P_{L'_1}} 
    P(X^{(1)} < t) \nonumber \\
    &+ \frac{\lambda_2 P_{L'_2}}{\lambda_2 P_{L'_2} + \lambda_1 
    P_{L'_1}} P(X^{(2)} < t) \label{eq:overall_proportion_within_target}
\end{align}

Here \(P(X_{(u,v)}^{(1)})\) and \(P(X_{(u,v)}^{(2)})\) are defined as the
proportion of individuals within the time target \(t\) when starting from state 
\((u,v)\).
These expression can be calculated by:

\begin{equation}\label{eq:proportion_within_target_type_1_from_state}
    P(X_{(u,v)}^{(1)} < t) = 
    \begin{cases}
        1 - \sum_{i=0}^{v-1} \frac{1}{i!} e^{-\mu t} (\mu t)^i, 
            & \textbf{if } C = 1 \\
            & \textbf{and } v>1 \\
        & \\
        1 - (\mu C)^{v-C} \mu  
            \sum_{k=1}^{\mid \vec{r} \mid} \sum_{l=1}^{r_k}
            \frac{\Psi_{k,l}(-\lambda_k)t^{r_k - l} 
            e^{-\lambda_k t}}{(r_k - l)! (l - 1)!},
            & \textbf{if } C > 1 \\
        \textbf{where } \vec{r}=(v - C, 1) \textbf{ and } 
            \vec{\lambda}=(C \mu, \mu) & \textbf{and } v > C \\
            \hspace{1.15cm} \lambda_0 = 0, r_0 = 1 \\
        & \\
        1 - e^{-\mu t},  & \textbf{if } v \leq C
    \end{cases}
\end{equation}

\begin{equation}\label{eq:proportion_within_target_type_2_from_state}
    P(X_{(u,v)}^{(2)} < t) = 
    \begin{cases}
        1 - \sum_{i=0}^{\min(v,T)-1} \frac{1}{i!} e^{-\mu t} (\mu t)^i,  
            & \textbf{if } C = 1 \\ 
            & \textbf{and } v, T > 1 \\
            & \\
        1 - (\mu C) ^ {\min(v,T) - C} \mu  
            &  \\
        \qquad \times \sum_{k=1}^{\mid \vec{r} \mid} \sum_{l=1}^{r_k} 
        \frac{\Psi_{k,l}(-\lambda_k)t^{r_k - l} 
        e^{-\lambda_k t}}{(r_k - l)! (l - 1)!}, 
            & \textbf{if } C > 1 \\
        \textbf{where } \vec{r}=(\min(v, T) - C, 1) 
            & \textbf{and } v, T  > C\\
        \hspace{1.15cm} \vec{\lambda}=(C \mu, \mu) \\
        \hspace{1.15cm} \lambda_0 = 0, r_0 = 1 \\
            & \\
        1 - e^{-\mu t}, & \textbf{if } v \leq C \\ 
            & \textbf{or } T \leq C \\
    \end{cases}
\end{equation}

The function \(\Psi_{k,l}\) used in equations 
(\ref{eq:proportion_within_target_type_1_from_state}) and 
(\ref{eq:proportion_within_target_type_2_from_state}) is defined as:

\begin{equation}
    \Psi_{k,l}(t) = 
    \begin{cases} 
        \frac{(-1)^{l} (l-1)!}{\lambda_2} \left[\frac{1}{t^l} - \frac{1}
        {(t + \lambda_2)^l}\right] , & k=1 \\
        - \frac{1}{t (t + \lambda_1)^{r_1}}, & k=2
    \end{cases} \nonumber \\
    % & \textbf{and} \quad \qquad \lambda_0 = 0, r_0 = 1
\end{equation}

Refer to appendix \ref{sec:appendix_mean_proportion} for a more in-depth 
explanation of the origins of equations 
(\ref{eq:proportion_within_target_type_1}) - 
(\ref{eq:proportion_within_target_type_2_from_state}).

\subsubsection{Example}
Figures \ref{fig:markov_vs_des_proportion_comparison_1},
\ref{fig:markov_vs_des_proportion_comparison_2} and 
\ref{fig:markov_vs_des_proportion_comparison_overall} show a comparison of the
mean proportion of individuals within target when using Markov chains and 
discrete event simulation.
The simulation was ran 100 times and the recorded proportions at each iteration 
is used to populate the violin plots.

\begin{figure}[H]
    \centering
    \includegraphics[width=.8\textwidth]{imgs/proportion_within_target_comparison/proportion_1.pdf}
    \caption{
        Comparison of proportion within target time for type 1 individuals 
        between values obtained from the Markov chain formulas and values 
        obtained from simulation. 
    }
    \label{fig:markov_vs_des_proportion_comparison_1}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.8\textwidth]{imgs/proportion_within_target_comparison/proportion_2.pdf}
    \caption{
        Comparison of proportion within target time for type 2 individuals 
        between values obtained from the Markov chain formulas and values 
        obtained from simulation. 
    }
    \label{fig:markov_vs_des_proportion_comparison_2}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.8\textwidth]{imgs/proportion_within_target_comparison/proportion_overall.pdf}
    \caption{
        Comparison of proportion within target time for both type of 
        individuals between values obtained from the Markov chain formulas 
        and values obtained from simulation. 
    }
    \label{fig:markov_vs_des_proportion_comparison_overall}
\end{figure}