### 2021-08-17 sections/appendix/descrete_event_simulation/main.tex
Apply changes based on Alex's suggestions

\section{Discrete Event Simulation}\label{sec:appendix_des}

For the purposes of this study, a discrete event simulation (DES) model was
constructed to support the Markov chain version described in 
section~\ref{sec:queueing_model}.
The queueing model was built in python using the Ciw 
library~\cite{ciwpython}.

The constructed model simulates a queueing system with two waiting spaces and 
two types of individuals. 
The expected behaviour of the nodes in Ciw have been modified such that 
individuals moving from waiting zone 2 into waiting zone 1 get blocked 
if there are more than \(T\) individuals in waiting zone 1.

The same performance measures described in sections 
\ref{sec:waiting_time}, \ref{sec:blocking_time} and 
\ref{sec:proportion_within_target} can also be calculated using the DES model.
The simulation can be ran a number of times to eliminate stochasticity and the
outcomes of the two methods can be directly comparable. 
% TODO Add more documentation (instructions on how to use code)

% TODO: Archive simulation python code