### 2021-09-16 sections/appendix/descrete_event_simulation/main.tex
Add Reference and Explanation section

\section{Discrete Event Simulation}\label{sec:appendix_des}

For the purposes of this study, a discrete event simulation (DES) model was
constructed to support the Markov chain version described in 
section~\ref{sec:queueing_model}.
The queueing model was built in python using the Ciw 
library~\cite{ciwpython}.

The constructed model simulates a queueing system with two waiting spaces and 
two types of individuals. 
The expected behaviour of the nodes in Ciw have been modified such that 
individuals moving from waiting zone 2 into waiting zone 1 get blocked 
if there are more than \(T\) individuals in waiting zone 1.

The same performance measures described in sections 
\ref{sec:waiting_time}, \ref{sec:blocking_time} and 
\ref{sec:proportion_within_target} can also be calculated using the DES model.
The simulation can be ran a number of times to eliminate stochasticity and the
outcomes of the two methods can be directly comparable.

\subsection{Tutorial: building the DES model}

The DES model is constructed in a generic way that so that it can be used for
any queueing system with two waiting spaces and two types of individuals.
For instance, consider a queueing system with the following parameters, as 
described in section~\ref{sec:queueing_model}:

\begin{multicols}{4}
    \begin{itemize}
        \item \( \lambda_1 = 2 \)
        \item \( \lambda_2 = 3 \)
        \columnbreak
        \item \( \mu = 1 \)
        \item \( C = 6 \)
        \columnbreak
        \item \( T = 10 \)
        \item \( N = 20 \)
        \columnbreak
        \item \( M = 10 \)
    \end{itemize}
\end{multicols}

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[>=stealth, scale=0.8],
        % the rectangle of parking queue
        \draw (0,0) -- ++(3cm,0) -- ++(0,-1.5cm) -- ++(-3cm,0);
        % the vertical lines of parking queue
        \foreach \i in {1,...,4, 7.1}
            \draw (3cm-\i*12pt,0) -- +(0,-1.5cm);

        % The label above Queue 1 -> M
        \node[anchor=north] at (1.5cm, 1cm) {\( 
            \overbrace{\qquad \qquad \qquad \qquad}^{10} 
        \)};

        % the rectangle in Queue 2
        \draw (5,1.25) -- ++(2cm,0) -- ++(0,-1.5cm) -- ++(-2cm,0);
        % the vertical lines in Queue 2
        \foreach \i in {1,...,4, 5.7}
            \draw (7cm-\i*10pt,1.25) -- +(0,-1.5cm);
        % The two vertical lines at the start of Queue 2 
        \draw (7cm-54pt,1.2) -- +(0,-0.5cm);
        \draw (7cm-54pt,0.3) -- +(0,-0.5cm);        
        \draw (7cm-51pt,1.1) -- +(0,-0.4cm);
        \draw (7cm-51pt,0.3) -- +(0,-0.4cm);

        % The label between the lines for T
        \node[anchor=north] at (5.19, 0.78 cm) {\small{\( 10 \)}};

        % The label above Queue 2 -> N
        \node[anchor=north] at (6.6cm, 2.2cm) {\(
            \overbrace{\qquad \qquad \qquad \qquad}^{20}
        \)};

        % the circle in Queue 2
        \draw (7.75,0.5) circle [radius=0.75cm] node {\(6\)};

        % Arrow line from Queue 2 outside
        \draw[->] (8.5,0.525) -- +(20pt,0);

        % Line from lambda_2 to Queue 1
        \draw[<-] (-0.3,-0.75) -- +(-40pt,0) node[left] {\( 3 \)};

        % Lines after Parking queue
        \draw[-] (3,-0.75) -- +(34pt,0);
        \draw (4.2, 0.525) -- (4.2, -0.75);

        % Lines for lambda_1 arrivals
        \draw (4.2, 1.8) -- +(-169.5pt,0) node[left] {\( 2 \)};
        \draw (4.2, 1.8) -- (4.2, 0.525);

        % Line from both individuals to Queue 2
        \draw[->] (4.2, 0.525) -- (4.8, 0.525);  
    \end{tikzpicture}
    \caption{A diagrammatic representation of the queueing model example}
    \label{fig:diagram_of_queueing_system_appendix_example}
\end{figure}

This model will be studied by using  
\lstinline[style=pystyle]{ambulance_game}.
Install the created library in your python environment, by running the 
following command in the command line:
\begin{lstlisting}[style=terminalstyle]
    $ pip install ambulance_game
\end{lstlisting}

Having installed the package, the following code can be used to simulate the 
queueing system defined earlier and get all the data records for a single run.

\begin{lstlisting}[style=pystyle]

    >>> import ambulance_game as abg
    >>> Simulation = abg.simulation.simulate_model(
    ...     lambda_1=3,
    ...     lambda_2=2,
    ...     mu=1,
    ...     num_of_servers=6,
    ...     threshold=10,
    ...     system_capacity=20,
    ...     buffer_capacity=10,
    ... )
    >>> all_records = Simulation.get_all_records()
    >>> all_records[3]
    Record(id_number=1, customer_class=0, node=2, arrival_date=0.4728763843239206, waiting_time=0.0, service_start_date=0.4728763843239206, service_time=0.5457131455415929, service_end_date=1.0185895298655134, time_blocked=0.0, exit_date=1.0185895298655134, destination=-1, queue_size_at_arrival=0, queue_size_at_departure=4)
\end{lstlisting}

The above block code outputs the fourth individual record from the simulation
object.
The simulation object can be used to view the every event that occurred in the 
simulation 
The data records can then be used to get overall performance measures about the
constructed queueing model.
The overall waiting time that individuals wait in waiting zone 1 can be acquired 
by running:

\begin{lstlisting}[style=pystyle]
    >>> import numpy as np
    >>> mean_wait = np.mean(
    ...    [record.waiting_time for record in all_records if record.node == 2]
    ... )
    0.2246928220731305
\end{lstlisting}

This value is the average waiting time of all the customers in the system for a
single run. 
By nature, discrete event simulation can output different results for different
runs of the same set of parameters.
This stochasticity can be reduced by running the simulation multiple times 
and then getting the mean waiting time from all the runs.

\begin{lstlisting}[style=pystyle]
    >>> all_simulations = abg.simulation.get_multiple_runs_results(
    ...     lambda_1=3,
    ...     lambda_2=2,
    ...     mu=1,
    ...     num_of_servers=6,
    ...     threshold=10,
    ...     system_capacity=20,
    ...     buffer_capacity=10,
    ...     seed_num=0,
    ...     num_of_trials=10,
    ... )
    >>> mean_wait = np.mean(
    ...     [np.mean(w.waiting_times) for w in all_simulations]
    ... )
    0.35037759808989827
\end{lstlisting}

\subsection{How-to guide}

The package can be installed by either running:

\begin{lstlisting}[style=terminalstyle]
    $ pip install ambulance_game
\end{lstlisting}

in the command line or via the instructions provided in the GitHub repository.
% TODO: Cite repository
The required arguments that need to be passed to the simulate\_model() function
are the following:
\begin{itemize}
    \item lambda\_1 (\(\lambda_1\)): The arrival rate of class 1 individuals.
    \item lambda\_2 (\(\lambda_2\)): The arrival rate of class 2 individuals.
    \item mu (\(\mu\)): The service rate of the servers.
    \item num\_of\_servers (\(C\)): The number of servers in the system.
    \item threshold (\(T\)): The threshold that indicates when to start blocking 
    class 2 individuals.
\end{itemize}

To get the simulation object with all the data records, the following code can be
used:
\begin{lstlisting}[style=pystyle]
    >>> import ambulance_game as abg
    >>> Simulation = abg.simulation.simulate_model(
    ...     lambda_1=3,
    ...     lambda_2=2,
    ...     mu=1,
    ...     num_of_servers=6,
    ...     threshold=10,
    ... )
    >>> Simulation.get_all_records()[1]
    Record(id_number=1, customer_class=0, node=2, arrival_date=0.5391449803891563, waiting_time=0.0, service_start_date=0.5391449803891563, service_time=0.17003700443114755, service_end_date=0.7091819848203038, time_blocked=0.0, exit_date=0.7091819848203038, destination=-1, queue_size_at_arrival=0, queue_size_at_departure=0)
\end{lstlisting}

Additional arguments that can be passed to the function are:
\begin{itemize}
    \item \textit{system\_capacity}: The maximum number of individuals in 
    waiting zone 2.
    \item \textit{buffer\_capacity}: The maximum number of individuals in 
    waiting zone 1.
    \item \textit{seed\_num}: The seed number for the random number generator.
    \item \textit{runtime}: How long to run the simulation for.
\end{itemize}

From a single run of the simulation the data records can be used the get the
average for certain performance measures. 
The following code can be used to get the mean waiting time, blocking time, 
service time and the proportion of individuals within target.

\begin{lstlisting}[style=pystyle]
    >>> records = Simulation.get_all_records()
    >>> mean_wait = np.mean(
    ...     [w.waiting_time for w in records]
    ... )
    >>> mean_block = np.mean(
    ...     [b.time_blocked for b in records]
    ... )
    >>> mean_service = np.mean(
    ...     [s.service_time for s in records]
    ... )
    >>> target = 1
    >>> proportion_within_target = np.mean(
    ...     [w.waiting_time <= target for w in records]
    ... )

\end{lstlisting}

To eliminate any stochasticity in the simulation, the simulation can be run
numerous times and get the average performance measures out of all the runs.

\begin{lstlisting}[style=pystyle]
    >>> import numpy as np
    >>> import ambulance_game as abg
    >>> all_simulations = abg.simulation.get_multiple_runs_results(
    ...     lambda_1=3,
    ...     lambda_2=2,
    ...     mu=1,
    ...     num_of_servers=6,
    ...     threshold=10,
    ...     system_capacity=20,
    ...     buffer_capacity=10,
    ...     seed_num=0,
    ...     runtime=2000,
    ...     num_of_trials=10,
    ... )
    >>> mean_wait = np.mean([
    ...     np.mean(w.waiting_times) for w in mean_simulations
    ... ])
    >>> mean_service = np.mean([
    ...     np.mean(s.service_times) for s in mult_results
    ... ])
    >>> mean_block = np.mean([
    ...     np.mean(b.blocking_times) for b in mult_results
    ... ])
\end{lstlisting}

To get the average proportion of individuals within target, run the following
code:

\begin{lstlisting}[style=pystyle]
    >>> import ambulance_game as abg
    >>> (
    ...     overall_proportion,
    ...     class_1_proportion,
    ...     class_2_proportion,
    ... ) = abg.simulation.
    get_mean_proportion_of_individuals_within_target_for_multiple_runs(    
    ...     lambda_1=3,
    ...     lambda_2=2,
    ...     mu=1,
    ...     num_of_servers=6,
    ...     threshold=10,
    ...     system_capacity=20,
    ...     buffer_capacity=10,
    ...     seed_num=0,
    ...     num_of_trials=10,
    ...     runtime=2000,
    ...     target=1,
    ... )
\end{lstlisting}
% TODO: Fix code for proportion of individuals within time target

\subsection{Reference}
The primary tool that was used in the construction of the discrete event 
simulation model was the python library \texttt{ciw}.
See Ciw's documentation for a more detailed explanation of how it works and 
what are its capabilities~\cite{ciwpython}.

Find below a detailed list of the functions that were created for the simulation
model:

\begin{itemize}
    \item \lstinline[style=pystyle]{build_model}: \textit{Builds a ciw object 
    that represents a model of a queuing network with two service centres.}
    \item \lstinline[style=pystyle]{build_custom_node}: \textit{Build a custom 
    node to replace the default ciw.Node. Inherits from the original ciw.Node 
    class and replaces two methods.}
    \item \lstinline[style=pystyle]{simulate_model}: \textit{Simulate the model
    by using the custom node and returning the simulation object}
    \item \lstinline[style=pystyle]{get_simulated_state_probabilities}: \textit{
    Calculates the vector pi in a dictionary format or an array format}
    \item \lstinline[style=pystyle]{extract_times_from_records}: \textit{Get the
    required times (waiting, service, blocking) out of ciw's records where all 
    individuals are treated the same way}
    \item \lstinline[style=pystyle]{extract_times_from_individuals}: \textit{
    Extract waiting times and service times for all individuals and proceed to 
    extract blocking times for just class 2 individuals}
    \item \lstinline[style=pystyle]{get_list_of_results}: \textit{Modify the 
    outputs so that they are returned in a list format where it is sometimes 
    easier to be used by other functions.}
    \item \lstinline[style=pystyle]{get_multiple_runs_results}: \textit{Get the 
    waiting times, service times and blocking times for multiple runs of the 
    simulation}
    \item \lstinline[style=pystyle]{get_average_simulated_state_probabilities}:
    \textit{This function runs \hfill\break
    get\_simulated\_state\_probabilities for multiple iterations to eliminate 
    any stochasticity from the results}
    \item \lstinline[style=pystyle]
    {get_mean_blocking_difference_between_two_systems}: \textit{Given a 
    predefined proportion of class's 2 arrival rate calculate the mean 
    difference between blocking times of two systems with a given set of 
    parameters}
    \item \lstinline[style=pystyle]{calculate_class_2_individuals_best_response}
    \textit{Obtains the optimal distribution of class 2 individuals such that 
    the blocking times in the two systems are identical and thus minimised}
    \item \lstinline[style=pystyle]
    {extract_total_individuals_and_the_ones_within_target_for_both_classes}:
    \hfill\break \textit{Extract the total number of individuals that pass 
    through the model and the total number of individuals that exit the model 
    within the given target.}
    \item \lstinline[style=pystyle]
    {get_mean_proportion_of_individuals_within_target_for_multiple_runs}: 
    \textit{Get the average proportion of individuals within target by running 
    the simulation multiple times}
\end{itemize}

\subsection{Explanation}

Based on Ciw's functionality the simulation model stores all data records in a
Record object. 
For every event that takes place a record is created with all the relevant
information.
For this specific library, the records that are stored, along with the 
range of values that they can take are as follows:

\begin{multicols}{2}
    \begin{itemize}
        \item \textbf{id\_number} \(\in R\).
        \item \textbf{customer\_class} \(= 0\)
        \item \textbf{node} \(= \{0, 1, 2, -1 \} \)
        \item \textbf{arrival\_date} \( \in R^+ \)
        \item \textbf{waiting\_time} \( \in R^+ \)
        \item \textbf{service\_start\_date} \( \in R^+ \)
        \item \textbf{service\_time} \( \in R^+ \)
        \item \textbf{service\_end\_time} \( \in R^+ \)
        \item \textbf{time\_blocked} \( \in R^+ \)
        \item \textbf{exit\_date} \( \in R^+ \)
        \item \textbf{destination} \( = \{1, 2, -1\} \)
        \item \textbf{queue\_size\_at\_arrival} \( \in N \)
        \item \textbf{queue\_size\_at\_departure} \( \in N \)
    \end{itemize}
\end{multicols}

% TODO: Archive simulation python code