### 2021-10-04 title.tex
Update author names

\title{A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface}
\author{Michalis Panayides, Vince Knight, Paul Harper}
\date{}