### 2021-11-11 sections/appendix/additional_figures/main.tex
Additional changes based on review
- Reword a sentence
- Move some plots from the main body to the appendix

\section{Type 1 and type 2 performance measure comparisons using simulation and
Markov chains}\label{sec:appendix_additional_figures}

\begin{figure}[H]
    \centering
    \includegraphics[width=.7\textwidth]{imgs/waiting_time_comparison/waiting_1.pdf}
    \caption{
        Comparison of mean waiting time for type 1 individuals between values 
        obtained from the Markov chain formulas and values obtained from 
        simulation.
    }
    \label{fig:markov_vs_des_waiting_time_comparison_1}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.7\textwidth]{imgs/waiting_time_comparison/waiting_2.pdf}
    \caption{
        Comparison of mean waiting time for type 2 individuals between values 
        obtained from the Markov chain formulas and values obtained from 
        simulation.
    }
    \label{fig:markov_vs_des_waiting_time_comparison_2}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.7\textwidth]{imgs/proportion_within_target_comparison/proportion_1.pdf}
    \caption{
        Comparison of proportion within target time for type 1 individuals 
        between values obtained from the Markov chain formulas and values 
        obtained from simulation. 
    }
    \label{fig:markov_vs_des_proportion_comparison_1}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.7\textwidth]{imgs/proportion_within_target_comparison/proportion_2.pdf}
    \caption{
        Comparison of proportion within target time for type 2 individuals 
        between values obtained from the Markov chain formulas and values 
        obtained from simulation. 
    }
    \label{fig:markov_vs_des_proportion_comparison_2}
\end{figure}