### 2021-11-19 title.tex
Add changes suggested by guide for authors
- Add abstract
- Add acknowledgements
- Fix the title page
- Change bibliogrpahy style

\title{A game theoretic model of the behavioural gaming that takes place at the 
EMS - ED interface\tnoteref{t1}}
\tnotetext[t1]{This document is the results of the research project funded by 
the heathcare improvement studies institute.}

\author[1]{Michalis Panayides\corref{cor1}}
\ead{panayidesM@cardiff.ac.uk}

\author[2]{Vince Knight}
\ead{knightva@cardiff.ac.uk}

\author[3]{Paul Harper}
\ead{harper@cardiff.ac.uk}

\address[1]{The Abacws 2.67, Senghennydd Rd, Cardiff CF24 4AX, United Kingdom}
\address[2]{The Abacws 2.52, Senghennydd Rd, Cardiff CF24 4AX, United Kingdom}
\address[3]{The Abacws 2.56, Senghennydd Rd, Cardiff CF24 4AX, United Kingdom}

\cortext[cor1]{Corresponding author}