### 2021-11-22 title.tex

\title{A game theoretic model of the behavioural gaming that takes place at the 
EMS - ED interface\tnoteref{t1}}
\tnotetext[t1]{This document is the results of the research project funded by 
the healthcare improvement studies institute.}

\author[1]{Michalis Panayides\corref{cor1}}
\ead{panayidesM@cardiff.ac.uk}

\author[1]{Vince Knight}
\ead{knightva@cardiff.ac.uk}

\author[1]{Paul Harper}
\ead{harper@cardiff.ac.uk}

\address[1]{School of Mathematics, Cardiff University, Cardiff CF24 4AX, United Kingdom}

\cortext[cor1]{Corresponding author}