### 2021-11-25 title.tex
Final changes before submitting paper

\title{A game theoretic model of the behavioural gaming that takes place at the 
EMS - ED interface\tnoteref{t1}}
\tnotetext[t1]{This document is the results of the research project funded by 
The Healthcare Improvement Studies (THIS) institute.}

\author{Michalis Panayides\corref{cor1}}
\ead{panayidesM@cardiff.ac.uk}

\author{Vince Knight}
\ead{knightva@cardiff.ac.uk}

\author{Paul Harper}
\ead{harper@cardiff.ac.uk}

\address{School of Mathematics, Cardiff University, Cardiff CF24 4AX, United Kingdom}

\cortext[cor1]{Corresponding author}