### 2022-02-21 sections/model_overview/main.tex
Change vocabulary in game section

\section{Overview of game theoretic model} \label{sec:model_overview}

The problem studied is a 3-player normal form game. The players are:

\begin{itemize}
    \item the decision makers of two Emergency Departments (EDs);
    \item the Emergency Medical Services (EMS) that distribute individuals in ambulances to the EDs.
\end{itemize}

This is a standard normal form game~\cite{Maschler2013},  
in that each player in this game has their own objectives which they aim to 
optimise.
More specifically, the EDs' objective is captured by an upper bound
of the time that a fixed proportion of individuals spend in the system, 
while the EMS aims to minimise the time that its ambulances 
are blocked.
This can be generalised for any such system where instead of EDs there are some
queueing systems and instead of the EMS there is some distributor that allocates individuals to the queueing systems.

The queueing systems of the hospitals are designed in such a way where they can
accept two types
of individuals. 
Type 2 individuals are the ones arriving in ambulances that are allocated by
the ambulance service and 
type 1 individuals from other sources (e.g walk-in patients, urgent patients).
Each hospital may then choose to block the individuals that arrive in
ambulances when the hospital reaches a certain capacity. 
The strategy sets for each hospital is the set 
\( \{T \in \mathbb{N} \;|\; 1 \leq T \leq N\} \) where \(N \in\{N_A, N_B\}\) are 
the total capacities of hospitals \(A\) and \(B\). We denote the chosen actions 
from the strategy set as \(T_A, T_B\) and call these \textit{threshold}s.

Both hospitals follow a queueing model with two waiting zones for 
individuals. 
The first waiting zone (i.e. the waiting space of the hospital) is where the 
patients queue right before receiving 
their service and has a capacity of \( N - C \), where \(N\) is the total 
capacity of the waiting space and \(C\) is the number of servers. 
% TODO: Mention what we mean by servers here?
The second waiting zone (i.e. the parking space for ambulances) is where 
ambulances, that are sent from the
EMS, stay until they are allowed to enter the first waiting zone.
The second waiting zone has a capacity of \(M\) and no servers.

This is shown diagrammatically in Figure ~\ref{fig:diagram_of_queueing_system}.

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[>=stealth, scale=0.8],
        % the rectangle of Queue 1
        \draw (0,0) -- ++(3cm,0) -- ++(0,-1.5cm) -- ++(-3cm,0);
        % The label above Queue 1 -> M
        \node[anchor=north] at (1.5cm, 1cm) {\( 
            \overbrace{\qquad \qquad \qquad \qquad}^{M} 
        \)};
        % The label below Queue 1 -> Waiting zone 2
        \node[anchor=north] at (1.5cm, -1.7cm) {\small{Waiting zone 2}};

        % the vertical lines in Queue 1
        \foreach \i in {1,...,5, 6.6}
        \draw (3cm-\i*13pt,0) -- +(0,-1.5cm);

        % % the circle in Queue 1
        % \draw (2.75,-0.75cm) circle [radius=0.75cm] node {\(0\)};

        % the rectangle in Queue 2
        \draw (5,1.25) -- ++(2cm,0) -- ++(0,-1.5cm) -- ++(-2cm,0);
        % the vertical lines in Queue 2
        \foreach \i in {1,...,4, 5.7}
        \draw (7cm-\i*10pt,1.25) -- +(0,-1.5cm);
        % The two vertical lines at the very start of Queue 2 
        \draw (7cm-54pt,1.2) -- +(0,-0.5cm);
        \draw (7cm-54pt,0.3) -- +(0,-0.5cm);        
        \draw (7cm-51pt,1.1) -- +(0,-0.4cm);
        \draw (7cm-51pt,0.3) -- +(0,-0.4cm);

        % The label between the lines for T
        \node[anchor=north] at (5.15, 0.77 cm) {\small{\( T \)}};

        % The label above Queue 2 -> N
        \node[anchor=north] at (6.6cm, 2.2cm) {\( 
            \overbrace{\qquad \qquad \qquad \qquad}^{N} 
        \)};
        % The label below Queue 2 -> Waiting zone 1
        \node[anchor=north] at (6.2cm, -0.5cm) {\small{Waiting zone 1}};

        % the circle in Queue 2
        \draw (7.75,0.5) circle [radius=0.75cm] node {\(C\)};

        % Arrow line from Queue 2 outside
        \draw[->] (8.5,0.525) -- +(20pt,0);

        % Line from lambda_2 to Queue 1
        \draw[<-] (-0.3,-0.75) -- +(-40pt,0) node[left] {\( \lambda_2 \)};
        % First line (horizontal) after Queue 1
        \draw[-] (3,-0.75) -- +(34pt,0);
        % Second line (vertical) after Queue 1
        \draw (4.2, 0.525) -- (4.2, -0.75);

        % First line (horizontal) from lambda_1
        \draw (4.2, 1.8) -- +(-169.5pt,0) node[left] {\( \lambda_1 \)};
        % Second line (vertical) from lambda_1
        \draw (4.2, 1.8) -- (4.2, 0.525);
        % Arrow line to Queue 2
        \draw[->] (4.2, 0.525) -- (4.8, 0.525);  
    \end{tikzpicture}
    \caption{A diagrammatic representation of the queueing model. 
    The threshold \(T\) only applies to type 2 individuals.
    If the number of individuals in waiting zone 2 is \(T\), only 
    individuals of the first type 
    are accepted (at a rate \(\lambda_1\)) and individuals of the second type 
    (arriving at a rate \(\lambda_2\)) are blocked in waiting zone 2.}
    \label{fig:diagram_of_queueing_system}
\end{figure}

Note here that both types of individuals can become lost to the system. 
An individual allocated from the ambulance service becomes lost to the system 
whenever 
an arrival occurs and the waiting zone 2 is at full capacity (\(M\) 
ambulances already parked).
Similarly, individuals get lost whenever they arrive at the waiting 
zone 1 and it is at full capacity (\(N - C\) individuals already waiting).

Following this queuing model, the two queueing systems' choice of strategy will 
then rely solely on satisfying their own 
objective, which is to satisfy:

\begin{equation}
    P(W < t) \geq \hat{P}
\end{equation}

where \(W\) is the waiting time of a potential individual, \(t\) is the time 
target and \(\hat{P}\) is the percentage of individuals need to be within that 
target. 
There are numerous objective functions that can be used to capture this 
behaviour. 
For example one approach is to use the threshold that maximises the probability 
that 
the mean waiting is more than the target time, and completely ignore the 
percentage goal.

\begin{equation}
    \arg \max_{T_i} \quad P(W_i < t)
\end{equation}

A more sophisticated objective function would be to get the proportion 
of individuals as close to the percentage aim. 
In other words, to find the threshold that minimises the difference between the 
probability and the percentage goal (or maximise its negation).

\begin{equation}\label{eq:obj_queueing_systems}
    \arg \max_{T_i} \quad -\left( \hat{P} - P(W_i < t) \right)^2
\end{equation}

The third player, the ambulance service, has their own choices to make and 
their own 
goals to satisfy.
The strategy set of the third player is the proportion \(0 \leq p_A \leq 1\) of 
individuals to send to hospital A.
Similarly the proportion of individuals to send to hospital B is given by
\(p_B = 1 - p_A\).
In addition, the ambulance service aims to minimise any potential blockages
that may occur, given the pair of thresholds chosen by the two hospitals.
Thus, its objective is to minimise the blocked time of the individuals 
(\(B_A\) and \(B_B\))
that they send to hospitals \(A\) and \(B\).
Apart from the time being blocked, an additional aspect that may affect the 
decision of the distributor is the proportion of lost individuals 
\(L_A\) and \(L_B\).
Equation \ref{eq:obj_distributor} can be used to capture a mixture 
between the two objectives \(L_i\) and \(B_i\) where \(i \in \{A, B\}\):
\begin{equation*}\label{eq:obj_distributor}
    (p_A, p_B) \quad s.t. \quad 
    \alpha L_A(p_A) + (1 - \alpha) B_A(p_A) = 
    \alpha L_B(p_B) + (1 - \alpha) B_B(p_B)
\end{equation*}

Here, \(\alpha\) represents the ``importance'' of each objective,
where high \(\alpha\) indicates a higher weight on the proportion of lost 
individuals and smaller \(\alpha\) a higher weight on the time blocked.

Using either equation (\ref{eq:obj_distributor}) or 
(\ref{eq:obj_queueing_systems}) gives an imperfect information extensive form 
game. 
An imperfect information game is defined as an extensive form game where some 
of the information about the game state is hidden for at least one of the 
players~\cite{Berwanger2008}. In this study the state of the problem that is
hidden is the threshold that each of the hospitals chooses to play.
In other words, each hospital chooses to play a strategy without 
knowing the other hospital's strategy.
The ambulance service then, fully aware of the chosen threshold strategies, 
distributes individuals among the two systems in order to minimise the time 
that its ambulances will be blocked. 
Figure \ref{fig:imperfect_info_game} illustrates this.

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[-, node distance = 3cm, scale=0.8]
        \node[anchor=north](HA){\(H_A\)};
        \node[anchor=north](HA_d1) at (3, 2){.};
        \node[anchor=north](HA_d2) at (3, -2){.};

        \path[->] (HA) edge node {}(HA_d1);
        \path[->] (HA) edge node {}(HA_d2);
        \path (HA_d1) edge [bend left] node {}(HA_d2);
        \path (HA_d1) [dashed] edge node {}(HA_d2);

        \node[anchor=north](HB) at (4.1, 0){\(H_B\)};
        \node[anchor=north](HB_d1) at (6.9, 2){.};
        \node[anchor=north](HB_d2) at (6.9, -2){.};

        \path[->] (HB) edge node {}(HB_d1);
        \path[->] (HB) edge node {}(HB_d2);
        \path(HB_d1) edge [bend left] node {}(HB_d2);

        \node[anchor=north](D) at (7.8, 0){\(D\)};
        \node[anchor=north](D_d1) at (10.8, 2){.};
        \node[anchor=north](D_d2) at (10.8, -2){.};

        \path[->] (D) edge node {}(D_d1);
        \path[->] (D) edge node {}(D_d2);
        \path(D_d1) edge [bend left] node {}(D_d2);
    \end{tikzpicture}
    \caption{Imperfect information Extensive Form Game between the distributor 
    and the 2 queueing systems}
    \label{fig:imperfect_info_game}
\end{figure}

The first 
queueing system \(H_A\) decides on a threshold, then the second system \(H_B\)
chooses its own threshold, without knowing the strategy of \(H_A\), and finally
the distributor makes its choice. Note here that the dotted line represents the
fact that \(H_B\) is unaware of the state of the game when making its own 
decisions.

In order to define the normal form game the two payoff matrices of the hospitals 
are required. From equation \ref{eq:obj_queueing_systems} the utilities of the
hospitals can be formulated as:

\begin{equation} \label{eq:payoff_entry}
    U_{T_A, T_B}^i = -\left( 
        \hat{P} - P(W_i < t) 
    \right)^2 \quad \textbf{where } i \in \{A, B\}
\end{equation}

Consequently, the payoff matrices of the game can be populated by these 
utilities:

\begin{equation} \label{eq:payoff_matrices}
    A = 
    \begin{pmatrix}
        U_{1,1}^A & U_{1,2}^A & \dots & U_{1,N_B}^A \\
        U_{2,1}^A & U_{2,2}^A & \dots & U_{2,N_B}^A \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^A & U_{N_A,2}^A & \dots & U_{N_A,N_B}^A \\
    \end{pmatrix}, \,
    B = 
    \begin{pmatrix}
        U_{1,1}^B & U_{1,2}^B & \dots & U_{1,N_B}^B \\
        U_{2,1}^B & U_{2,2}^B & \dots & U_{2,N_B}^B \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^B & U_{N_A,2}^B & \dots & U_{N_A,N_B}^B \\
    \end{pmatrix}
\end{equation}

Based on the choice of strategy of these two hospitals, the ambulance service 
will then 
make their own choice of the proportion of individuals to send to each system.